Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-10-23 experiments.tex
AA: ultimati i TODO AND sul testo
added platform description

\section{A Case Study: VLIW Hardware/Software Design Space}
\secL{ee}
\subsection{Motivation}
In order to evaluate the behavior of the proposed approach
we use a parameterized Very Long Instruction
Word (VLIW) architecture~\cite{kathail_tr00} as testbed. The philosophy
behind a VLIW system makes this a natural choice for several reasons:
\begin{itemize}
\item \emph{Multi-Objective trade-offs}: VLIW architectures allow designers
to trade between power, energy and performances objectives, resulting
in extended Pareto sets which are an ideal ground for our testing
purposes focused on parameter-space representation
\item \emph{Compiler awareness}: code scheduling for the execution of
applications is statically obtained by the compiler which
performs code transformations ``being aware'' of the underlying hardware
configuration.  This tight hardware/software dependence makes the
VLIW scenario perfectly suitable for investigating hardly predictable
parameter interactions.
\end{itemize}

@@

@@

%------------------------------------------------------------------------------
\subsection{Evaluation campaign}
In order to obtain the results presented in this section, we implemented our exploration strategy in Epic Explorer~\cite{ascia2003epic}, an open source framework for the design space explora=tion of embedded systems. \comment{AA: continuare descrivendo un po' il workflow. The reader can refer to~\cite{ascia2003epic} for more details about the estimation models of the power dissipation and execution time. Scrivere che it is possible to run benchmarks }.

>>>>

\section{A Case Study: VLIW Hardware/Software Design Space}
\secL{ee}
Since our main focus is the embedded system design, we apply PS to a
related case study.  In this section, we briefly describe the
parameterized platform used as testbed for the experiments, the
general evaluation flow along with the high-level estimation models
and applications used as benchmarks.

\subsection{Motivation}
In order to evaluate the behavior of the proposed approach
we use a parameterized Very Long Instruction
Word (VLIW) architecture~\cite{kathail_tr00} as testbed. The philosophy
behind a VLIW system makes this a natural choice for several reasons:
\begin{itemize}
\item \emph{Multi-Objective trade-offs}: VLIW architectures allow designers
to trade between power, energy and performances objectives, resulting
in extended Pareto sets which are an ideal ground for our testing
purposes focused on parameter-space representation
\item \emph{Compiler awareness}: code scheduling for the execution of
applications is statically obtained by the compiler which
performs code transformations ``being aware'' of the underlying hardware
configuration.  This tight hardware/software dependence makes the
VLIW scenario perfectly suitable for investigating hardly predictable
parameter interactions.
\end{itemize}

@@

%------------------------------------------------------------------------------
\subsection{Evaluation Flow}

To evaluate and compare the performance indexes of different
architectures for a specific application, one needs to simulate the
architecture running the code of the application. To make
architectural exploration possible both the compiler and the simulator
have to be retargetable. Trimaran~\cite{trimaran_hp} provides these
tools and thus represents the pillar around which we have constructed
\ee~\cite{palpatti_estimedia03}. \ee{} is a framework that not only
allows us to evaluate any instance of a platform in terms of area,
performance and power, but also implements various techniques for
exploration of the design space.

We will now show a functional scheme highlighting the main blocks of
the \ee{} framework and the interface with the Trimaran tools. The
input for the whole evaluation flow is the source of the application
involved in the optimization and the configuration of the architecture
being evaluated.  With reference to Figure~\ref{fig:evaluation_flow}
this input is represented by the blocks \emph{App.c} and
\emph{Config}.
\begin{figure}
        \centering
        \includegraphics[width=0.65\textwidth]{pictures/evaluation_flow.eps}
        \caption{Block diagram of the framework.}
        \label{fig:evaluation_flow}
\end{figure}

The application (App.c) is first compiled by the Trimaran's compiler
front-end (IMPACT). At the end of the compilation flow,
Trimaran supplies a simulation library (Emulib) which makes it
possible to execute the VLIW code produced by Elcor, generating a file
(Stats) containing the execution statistics (e.g., instruction mix,
execution cycles, utilization of functional units, etc.). A cache
simulator, along with a bus simulator, is used to gather information
about the behavior of the memory hierarchy in terms of miss rate and
data/address traffic on the interconnection buses.

Together with the configuration of the system, the statistics produced
by simulation contain all the information needed to apply the area,
performance and power consumption estimation model implemented in the
\emph{Estimator} component of \ee{}. The results obtained by these
models are the input for the \emph{Explorer} component. This component
executes an optimization algorithm, the aim of which is to modify the
parameters of the configuration so as to minimize the three cost
functions (area, execution time and energy/power consumption).

\subsection{Estimation Models}
Several models have been implemented in the \ee{} platform in order to
estimate the power/energy/performance indexes of each simulated system
configuration.

The average power consumed by the processor was estimated using an
adaptation of the Cai-Lim model~\cite{cai_micro99} to the VLIW
processor. As regards the cache subsystem, a transition-based model
was used, according to the equations described
in~\cite{kamble_islped97}.  The main memory energy is based on the
model in~\cite{shiue_dac99} and assumes a per main memory access
energy of $4.95 \times 10^{-9} J$ based on the data for the Cypress
CY7C1326-133 memory chip.  The contribution towards power consumption
made by the interconnection system was calculated by counting the
number of transitions on the bus lines and applying the formula
$P_{bus} = 1/2 V_{dd}^2 \alpha f C_l$ where $V_{dd}$ is the supply
voltage, $\alpha$ is the switching activity, $f$ is the clock
frequency and $C_l$ is the capacity of a bus line. For the on-chip
buses we also considered the coupling capacitances between bus lines,
using the model in~\cite{lekatsas_dac01}. For a detailed description
of the models used and their adaption to the case of a VLIW based
system see~\cite{palpatti_estimedia03}.

@@
